
# IMDB ratings: Differences between directors

I would like you to explore whether the mean IMDB rating for Steven Spielberg and Tim Burton are the same or not. I have already calculated the confidence intervals for the mean ratings of these two directors and as you can see they overlap. 

```{r load-movies-data}
movies <- read_csv(here::here("data", "movies.csv"))
glimpse(movies)
```

*Null Hypothesis:* Both Directors (Steven Spielberg & Tim Burton) have same average ratings on IMDB
*Alternative Hypothesis:* The two directors **do not** have the same average ratings.

```{r nullhypothesis}
movies_new <- movies %>% 
  filter(director %in% c("Steven Spielberg","Tim Burton")) 

t.test(rating ~ director, data = movies_new)

imdb_ratings <- movies_new %>%
  group_by(director) %>% 
  summarise(mean_rating = mean(rating),
            count = n(),
            margin_of_error = qt(0.975, count-1)*sd(rating)/sqrt(count),
            lower = mean_rating - margin_of_error,
            upper = mean_rating + margin_of_error)
```

The p value is 0.01 (less than 0.05), so we can reject the null hypothesis.

```{r fig.align="center",fig.width = 10}
ggplot(imdb_ratings,aes(x = mean_rating, y = reorder(director, desc(director)), colour = director)) + 
  geom_point(size = 5) +
  geom_errorbar(aes(xmin = lower, xmax = upper), width = 0.1, size = 2) +
  geom_rect(aes(xmin=max(lower), xmax=min(upper), ymin=0, ymax=Inf),
            color='grey', alpha=0.2) +
  
  geom_text(aes(label = round(mean_rating, digits = 2), x = mean_rating),
            size = 7, colour = "black", nudge_y = 0.15) + 
  geom_text(aes(label = round(lower, digits = 2), x = lower),
            size = 5, colour = "black",nudge_y = 0.1) +
  geom_text(aes(label = round(upper, digits = 2), x = upper),
            size = 5, colour = "black",nudge_y = 0.1) +
  
  labs(title = "Do Spielberg and Burton have the same mean IMDB ratings?",
       subtitle = "95% confidence intervals overlap" ,
       x = "Mean IMDB Rating", y = " ") +
  theme_bw() +
  theme(legend.position = "none")
```

The average rating for Steven Spielberg is higher than the average ratings for Tim Burton (at a 95% confidence interval). Thus we can reject the Null Hypothesis.

Using infer package:

```{r}
obs_diff_movies <- movies_new %>%
  specify(rating ~ director) %>%
  calculate(stat = "diff in means", order = c("Steven Spielberg", "Tim Burton"))

null_dist_movies <- movies_new %>%
  specify(rating ~ director) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 1000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("Steven Spielberg", "Tim Burton"))

null_dist_movies %>% visualize() +
  shade_p_value(obs_stat = obs_diff_movies, direction = "two-sided")

null_dist_movies %>%
  get_p_value(obs_stat = obs_diff_movies, direction = "two_sided")
```

Based on infer package as well, the p value is 0.02 (less than 0.5), so we can reject null hypothesis.
To conclude, Steven Spielberg and Tim Burton do not have the same average IMDB Ratings.

