

# Omega Group plc- Pay Discrimination

At the last board meeting of Omega Group Plc., the headquarters of a large multinational company, the issue was raised that women were being discriminated in the company, in the sense that the salaries were not the same for male and female executives. A quick analysis of a sample of 50 employees (of which 24 men and 26 women) revealed that the average salary for men was about 8,700 higher than for women. This seemed like a considerable difference, so it was decided that a further analysis of the company salaries was warranted. 

## Loading the data

```{r load_omega_data}
omega <- read_csv(here::here("data", "omega.csv"))
glimpse(omega) # examine the data frame
```

Summary statistics: 

## Relationship Salary - Gender ?

```{r, confint_single_valiables}
# Summary Statistics of salary by gender
mosaic::favstats (salary ~ gender, data=omega)

# Dataframe with two rows (male-female) and having as columns gender, mean, SD, sample size, 

gender_data_frame <- omega %>% 
  group_by(gender) %>% 
  summarise(mean = mean(salary),sd = sd(salary),size = n())


# the t-critical value, the standard error, the margin of error, 
gender_data_frame <- gender_data_frame %>% 
  mutate(tcritical = qt(p=0.05/2,df=size-1,lower.tail = FALSE),standard_error=sd/sqrt(size),margin_of_error=tcritical*standard_error)

# and the low/high endpoints of a 95% confidence interval

gender_data_frame <- gender_data_frame %>% 
  mutate(low=mean-tcritical*standard_error,high=mean+tcritical*standard_error)

print(gender_data_frame)

```

We can see that for women, the salary ranges from 61486 to 67599 while for men it ranges from 70088 to 76390 with 95% confidence. Since these ranges have no common point, it seems that there really is a significant difference between the salaries. 


Hypothesis testing: null hypothesis = the mean difference in salaries is zero, or on average, men and women make the same amount of money. 

```{r, hypothesis_testing1}
# hypothesis testing using t.test() 

t.test(salary~gender,data=omega)


# hypothesis testing using infer package
set.seed(1234)
mean_diff_df <- omega %>% 
  specify(salary~gender) %>% 
  hypothesize(null="independence") %>% 
  generate(reps=1000,type="permute") %>% 
  calculate(stat="diff in means",order = c("male","female"))

observed_statistic <- omega %>%
  specify(salary~gender) %>% 
  calculate(stat="diff in means",order = c("male","female"))

mean_diff_df %>%
  visualize()+
  shade_p_value(observed_statistic, direction = "two-sided")

mean_diff_df %>% 
  get_p_value(obs_stat = observed_statistic, direction = "two-sided")

  

```

Both t-tests show the same output, p-value is almost 0. That means the we should reject the null hypothesis which stated that salary of male and female workers were the same. There is definitely a difference between the means.

## Relationship Experience - Gender?

At the board meeting, someone raised the issue that there was indeed a substantial difference between male and female salaries, but that this was attributable to other reasons such as differences in experience. A questionnaire send out to the 50 executives in the sample reveals that the average experience of the men is approximately 21 years, whereas the women only have about 7 years experience on average (see table below).

```{r, experience_stats}
# Summary Statistics of salary by gender
favstats (experience ~ gender, data=omega)

```


```{r, hypothesis_testing2}
# hypothesis testing using t.test() 

t.test(experience~gender,data=omega)

# hypothesis testing using infer package
set.seed(1234)
mean_diff_exp_df <- omega %>% 
  specify(experience~gender) %>% 
  hypothesize(null="independence") %>% 
  generate(reps=1000,type="permute") %>% 
  calculate(stat="diff in means",order = c("male","female"))

observed_statistic <- omega %>%
  specify(experience~gender) %>% 
  calculate(stat="diff in means",order = c("male","female"))

mean_diff_exp_df %>%
  visualize()+
  shade_p_value(observed_statistic, direction = "two-sided")

mean_diff_exp_df %>% 
  get_p_value(obs_stat = observed_statistic, direction = "two-sided")

```

There is also a significant difference between the experiences of males and females since p-value is almost zero. This conclusion endangers the previous conclusion that there was a difference between male and female salaries. But now we see that this difference can be caused by the experience of workers.

## Relationship Salary - Experience ?

Someone at the meeting argues that clearly, a more thorough analysis of the relationship between salary and experience is required before any conclusion can be drawn about whether there is any gender-based salary discrimination in the company.

Analyse the relationship between salary and experience. Draw a scatterplot to visually inspect the data


```{r, salary_exp_scatter}
omega %>% 
  ggplot(aes(x=experience,y=salary,col=gender))+
  geom_point()


```


## Check correlations between the data

You can use `GGally:ggpairs()` to create a scatterplot and correlation matrix. Essentially, we change the order our variables will appear in and have the dependent variable (Y), salary, as last in our list. We then pipe the dataframe to `ggpairs()` with `aes` arguments to colour by `gender` and make ths plots somewhat transparent (`alpha  = 0.3`).

```{r, ggpairs}
omega %>% 
  select(gender, experience, salary) %>% #order variables they will appear in ggpairs()
  ggpairs(aes(colour=gender, alpha = 0.3))+
  theme_bw()
```

> Look at the salary vs experience scatterplot. What can you infer from this plot? Explain in a couple of sentences

The scatter plot and the correlation figure of 0.8 shows that salary is highly explained by the experience of a worker. Since there is significant difference between the experience of genders, the salary difference is explained by the experience.


