
# Challenge 1: Yield Curve inversion

```{r download_historical_yield_curve, warning=FALSE}
#Load data
yield_curve <- read_csv(here::here("data", "yield_curve.csv"))
```

## Yields on US rates by duration since 1960

```{r yield_curve_1, out.width="100%"}
#PLotting a graph of rates since 1960 
yield_curve %>% 
  ggplot(aes(x=date,y=value,color = duration))+
  geom_line(size=0.2)+
  
#Faceting the graphs by duration
facet_wrap(~ factor(duration, 
                        levels = c("3-Month Treasury Bill",
                                   "6-Month Treasury Bill",
                                   "1-Year Treasury Rate",
                                   "2-Year Treasury Rate",
                                   "3-Year Treasury Rate",
                                   "5-Year Treasury Rate",
                                   "7-Year Treasury Rate",
                                   "10-Year Treasury Rate",
                                   "20-Year Treasury Rate",
                                   "30-Year Treasury Rate")), 
               ncol = 2)+
  theme_bw()+
  labs(title="Yiels of U.S Treasury rates since 1960",color="")

```

## Monthly yields on US rates by duration since 1999 on a year-by-year basis

```{r yield_curve_2, out.width="100%"}

yield_curve %>%
  
  #Deleting the years previous to 1999
  filter(year(date)>=1999) %>% 
  
  #Plotting US rates by duration since 1999
  ggplot(aes(x=maturity,y=value,group= date,color=as.factor(year(date) )))+
  geom_line(size=0.2)+
  
  #Faceting by years
  facet_wrap(~year(date), ncol=4)+
  
  #Reordering the x axis
  scale_x_discrete(limits=c('3m','6m','1y','2y','3y','5y','7y','10y','20y','30y'))+
  theme_bw()+
  theme(legend.position = "none")+
  labs(title="US Yield Curve",x="Maturity",y="Yield %",caption=NULL)
 
```



### 3-month and 10-year yields since 1999

```{r yield_curve_3, echo=FALSE, out.width="100%"}

#PLotting the 3m and 10y rates since 1999
yield_curve %>% 
  
  #Selecting the 3m and 10y rates
  filter(maturity=="3m" | maturity=="10y") %>% 
  
  #Deleting years previous 1999
  filter(year(date)>=1999) %>% 
  
  ggplot(aes(x=date,y=value,color=duration),size=0.2)+
  geom_line()+
  theme_bw()+
  labs(title="Yield on 3-month and 10-year US Treasury rates since 1999",x=NULL,y="%",color="")
```


```{r setup_US-recessions, warning=FALSE}

# get US recession dates after 1946 from Wikipedia 
# https://en.wikipedia.org/wiki/List_of_recessions_in_the_United_States

recessions <- tibble(
  from = c("1948-11-01", "1953-07-01", "1957-08-01", "1960-04-01", "1969-12-01", "1973-11-01", "1980-01-01","1981-07-01", "1990-07-01", "2001-03-01", "2007-12-01","2020-02-01"),  
  to = c("1949-10-01", "1954-05-01", "1958-04-01", "1961-02-01", "1970-11-01", "1975-03-01", "1980-07-01", "1982-11-01", "1991-03-01", "2001-11-01", "2009-06-01", "2020-04-30") 
  )  %>% 
  mutate(From = ymd(from), 
         To=ymd(to),
         duration_days = To-From)
recessions
```

```{r setup_US-recessions_1, warning=FALSE}

#Plotting final graph

#Changing the format from long to wide in order to compute the spread and plot it
yield_wide <- yield_curve %>% 
  
  #deleting two columns that prevent the dataframe to be converting into long format
  select(-series_id) %>% 
  select(-duration) %>% 
  
  #Creating the columns for each maturity and assigning the respective yields 
  pivot_wider(names_from = "maturity",values_from = "value") %>% 
  
  #Computing the spread
  mutate(spread =`10y`-`3y`)

  #Deleting the receissions  data previous to 1960 
  recessions <- recessions %>% 
  filter(year(From)>=1960)

  #Creating the plot
  ggplot()+
  #Plotting the spread line
  geom_line(data=yield_wide,aes(x=date,y=spread),size=0.2)+

  #Plotting the blue and red ribbons
  geom_ribbon(data=yield_wide,aes(x=date, ymin = 0, ymax = spread, fill = spread > 0),alpha=0.2)+
  scale_fill_manual(values=c("red","blue"), name="fill")+

  #Plotting the grey shades
  geom_rect(data = data.frame(x1=recessions$From, x2=recessions$To, y1=-Inf, y2=Inf),aes(xmin=x1,xmax=x2,ymin=y1,ymax=y2),alpha=0.2)+
  #Deleting legend
  guides(fill=FALSE)+
  labs(title="Yield Curve Inversion: 10-year minus 3-month US. Tresury rates", subtitle="Difference in % points, monthly averages. Shaded areas correspond to recession",y="Difference(10 year- 3month) yield in %")+
  theme_bw()

```

A yield curve flattening means a 10y-3m spread near zero. On the previous graph, we can see that before the two regressions since 1999, the spread contracted to negative values (before the 2001 recession) and near-zero values (before the 2007-2009 recession) though with a steeper fall.The graph seems to corroborate the idea that a flattening curve is a signal for recession. What would be interesting to do is a regression between spread and recession (with lagged variable) to verify causality.

